\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-04-25 forwards-and-backwards.tex
fixups

\begin{theorem} \label{thm:lineages}
    In the context of Definition~\ref{def:lineage_generator},
    \begin{align}
        \label{eqn:lineage_generator}
        \Lgen_s f
        &=
        \frac{r}{\varphi_s}
        \left[
            \DG^*(\gamma \varphi_s f) 
            - f \DG^* (\gamma \varphi_s)
        \right] \\
        &= \label{eqn:lineage_generator2}
        r\gamma
        \left[
            \sum_{ij} \covq_{ij} \partial_{ij} f
            + 2 \sum_j \left(
               \sum_i \partial_i (C_{ij} \log(\gamma \varphi_s))
            \right) \partial_j f
            - \sum_j \meanq_j \partial_j f
        \right] .
    \end{align}
\end{theorem}

@@

To simplify to equation \eqref{eqn:lineage_generator2},
this, first define
$
    \DD f(x) = \sum_{ij} \covq_{ij} \partial_{ij} f(x),
$
% and note that $\DD$ satisfies the following identity:
% \begin{align}
%     \DD(fg)
%     &=
%     f \DD g
%     + 2 (\covq \grad f) \cdot \grad g
%     + g \DD f .
% \end{align}
and so the adjoint of $\DD$ is
$$
    \DD^* f(x)
    =
    \sum_{ij} \partial_{ij} (\covq_{ij} f(x)) .
$$
Note that $\DD^*$ satisfies the following identity:
\begin{align}
    \DD^*(fg)
    &=
    \sum_{ij} \left\{
        g \partial_{ij} (\covq_{ij} f)
        + 2 \partial_{i} (\covq_{ij} f) \partial_j(g)
        + \covq_{ij} f \partial_{ij} g
    \right\} \\
    &=
    g \DD^* f
    + 2 \mathcal{H} f \cdot \grad g
    + f \DD g ,
\end{align}
where $(\mathcal{H} f)_j = \sum_i \partial_i (C_{ij} f)$.
    \comment{(Need better notation for that.)}
So, with $f = \gamma \varphi_s$,
\begin{eqnarray*}
\hat{\mathcal L}_s g
    &=&
    \frac{r}{\varphi_s} \left[
        \DD^*(\gamma \varphi_s g) - \grad \cdot (\gamma \varphi_s g \meanq)
        - g \DD^*(\gamma \varphi_s) + g \grad \cdot (\gamma \varphi_s \meanq)
    \right] \\
    &=&
    \frac{r}{\varphi_s} \left[
        \gamma \varphi_s \DD g
        + 2 \mathcal{H} (\gamma \varphi_s) \cdot \grad g
        - \gamma \varphi_s \meanq \cdot \grad g
    \right] \\
    &=&
    r \gamma \left[
        \DD g
        + 2 (\mathcal{H} \log(\gamma \varphi_s)) \cdot \grad g
        - \meanq \cdot \grad g
    \right] ,
\end{eqnarray*}
in agreement with \eqref{eqn:heuristic_lineage_generator}.

@@

In this section, we need only make
simplify equation \eqref{eqn:lineage_generator}
in the case where $\covq = \sigma^2 I$.

@@

\begin{proof}[of Corollary~\ref{cor:wavefront}
    \comment{
        TODO make non-redundant with proof of Theorem \ref{thm:lineages.}.
    }

>>>>

\begin{theorem} \label{thm:lineages}
    In the context of Definition~\ref{def:lineage_generator},
    \begin{align}
        \label{eqn:lineage_generator}
        \Lgen_s f
        &=
        \frac{r}{\varphi_s}
        \left[
            \DG^*(\gamma \varphi_s f) 
            - f \DG^* (\gamma \varphi_s)
        \right] \\
        &= \label{eqn:lineage_generator2}
        r\gamma
        \left[
            \sum_{ij} \covq_{ij} \partial_{ij} f
            + 2 \sum_j \vec{m}_j \partial_j f
        \right] ,
    \end{align}
    where $\vec{m}$ is the vector
    $$
    \vec{m}_j
    =
    2 \sum_i C_{ij} \partial_i \log(\gamma \varphi_s)
    + 2 \sum_i \partial_i C_{ij}
    - \meanq_j .
    $$
\end{theorem}

@@

To simplify to equation \eqref{eqn:lineage_generator2},
this, first define
$
    \DD f(x) = \sum_{ij} \covq_{ij} \partial_{ij} f(x),
$
% and note that $\DD$ satisfies the following identity:
% \begin{align}
%     \DD(fg)
%     &=
%     f \DD g
%     + 2 (\covq \grad f) \cdot \grad g
%     + g \DD f .
% \end{align}
and so the adjoint of $\DD$ is
$$
    \DD^* f(x)
    =
    \sum_{ij} \partial_{ij} (\covq_{ij} f(x)) .
$$
Note that $\DD^*$ satisfies the following identity:
\begin{align*}
    \DD^*(fg)
    &=
    \sum_{ij} \left\{
        g \partial_{ij} (\covq_{ij} f)
        + 2 \partial_{i} (\covq_{ij} f) \partial_j(g)
        + \covq_{ij} f \partial_{ij} g
    \right\} \\
    &=
    \sum_{ij} \left\{
        g \partial_{ij} (\covq_{ij} f)
        + 2 f \partial_{i} (\covq_{ij}) \partial_j(g)
        + 2 \covq_{ij} \partial_{i} (f) \partial_j(g)
        + \covq_{ij} f \partial_{ij} g
    \right\} \\
    &=
    g \DD^* f
    + 2 f \vec{c} \cdot \grad g
    + 2 (C \grad f) \cdot \grad g
    + f \DD g ,
\end{align*}
where $\vec{c}_j = \sum_i \partial_j C_{ij}$.
So, with $f = \gamma \varphi_s$,
\begin{eqnarray*}
\hat{\mathcal L}_s g
    &=&
    \frac{r}{\varphi_s} \left[
        \DD^*(\gamma \varphi_s g) - \grad \cdot (\gamma \varphi_s g \meanq)
        - g \DD^*(\gamma \varphi_s) + g \grad \cdot (\gamma \varphi_s \meanq)
    \right] \\
    &=&
    \frac{r}{\varphi_s} \left[
        \gamma \varphi_s \DD g
        + 2 \gamma \varphi_s \vec{c} \cdot \grad g
        + 2 (C \grad (\gamma \varphi_s)) \cdot \grad g
        - \gamma \varphi_s \meanq \cdot \grad g
    \right] \\
    &=&
    r \gamma \left[
        \DD g
        + 2 \vec{c} \cdot \grad g
        + 2 (C \grad \log(\gamma \varphi_s)) \cdot \grad g
        - \meanq \cdot \grad g
    \right] ,
\end{eqnarray*}
which is equation \eqref{eqn:lineage_generator}.

@@

In this section, we need only
simplify equation \eqref{eqn:lineage_generator}
in the case where $\covq = \sigma^2 I$.

@@

\begin{proof}[of Corollary~\ref{cor:wavefront}
    \comment{
        TODO make non-redundant with proof of Theorem \ref{thm:lineages}.
    }
